### 2020-11-20 README.Rmd

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width = 16/2,
  fig.height = 9/2,
  message = FALSE,
  warning = FALSE,
  cache = TRUE
)
```

@@

This package is designed to work for spaitally mapped, repeat censused forests plots. The package has commands to fit models of tree growth based on neighborhood competition. This can be used to estimate species-specific competition coefficients. The models can be fit with a spatial cross validation scheme to avoid overfitting. Additionally these models can test whether the idenity of competitors matters, by shuffling competitor identity and testing whether model preformance changes. See Allen and Kim (2020) for a full descripition.

@@

```{r}
library(tidyverse)
library(forestecology)
library(sf)
library(sfheaders)
library(blockCV)
library(yardstick)
```

Here we provide an example with a small dateset. First we combine the two census data files into a single `tibble` with the growth of each surviving individual.

@@

# Filter out resprouts
census_df2_ex_no_r <- census_df2_ex %>%
  filter(!str_detect(codes, 'R'))

id <- 'ID'

ex_growth_df <-
  # Merge both censuses and compute growth:
  compute_growth(census_df1_ex, census_df2_ex_no_r, id) %>%
  mutate(
    sp = to_any_case(sp),
    sp = as.factor(sp)) 
```

All growth models in the package assume that two individuals compete if they are less than a specified distance away. Here we set this distance and put all individuals within this distance of a plot boundray in a buffer (since we do not have full information about their competitors).

```{r}
data("ex_study_region")

# set max dist
max_dist <- 1

# add buffer
ex_growth_df <- ex_growth_df %>%
  add_buffer_variable(direction = "in", size = max_dist, region = ex_study_region)

@@

We then set blocks for the cross validation scheme.

```{r}
# Bert's solution: Manually create a blocks sf object
fold1 <- rbind(c(0, 0), c(5, 0), c(5, 5), c(0, 5), c(0, 0))
fold2 <- rbind(c(5, 0), c(10, 0), c(10, 5), c(5, 5), c(5, 0))
blocks <- bind_rows(
  sf_polygon(fold1),
  sf_polygon(fold2) ) %>%
  mutate(foldID = c(1, 2))

# Plot
ggplot() +
  geom_sf(data = ex_growth_df, aes(col = buffer), size = 2) +
  geom_sf(data = blocks, fill = "transparent")

# fit spatialBlock()
ex_cv_grid <- spatialBlock(
  speciesData = ex_growth_df,
  verbose = FALSE,
  k = 2,
  # Note new arguments
  selection = "systematic",
  blocks = blocks#,
 # foldsCol = "foldID"
)

@@

# Visualize grid
ex_cv_grid$plots +
  geom_sf(data = ex_growth_df, aes(col = foldID), size = 2)

# Deliverable
ggplot() +
  geom_sf(data = ex_growth_df, aes(col = foldID, shape = buffer))

ex_cv_grid_sf <- ex_cv_grid$blocks %>%
  st_as_sf()
```

We then create a `focal_v_comp` tibble which has a row for each competing pair of individuals (each pair of individuals within `max_dist` of one another).

```{r}

# focal v comp
focal_vs_comp_ex <- ex_growth_df %>% 
  create_focal_vs_comp(max_dist, cv_grid_sf = ex_cv_grid_sf, id = "ID")
```

With `focal_v_comp` we can then run the growth model and get predictions from it. Here we run the model without cross validation. Posterior estimates can be plotted to give, for example, the lambda matrix of competition coefficents.

```{r}
# fit the model
posterior_param_ex <- focal_vs_comp_ex %>% 
  fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)

# get predictions
predictions <- focal_vs_comp_ex %>%
  predict_bayesian_model(posterior_param = posterior_param_ex) %>%
  right_join(ex_growth_df, by = c("focal_ID" = "ID"))
predictions %>%
  rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)

# plot posteriors
plot_ex <- posterior_param_ex %>%
  plot_posterior_parameters()

plot_ex[["lambda"]]

```

Here we repeat the model fitting but with cross-validation. Note the increase in RMSE.

```{r}
# run model with CV
ex_bw <- focal_vs_comp_ex %>% 
  run_cv(max_dist = max_dist, cv_grid = ex_cv_grid) %>%
  right_join(ex_growth_df, by = c("focal_ID" = "ID"))

ex_bw %>%
  rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)

```

>>>>

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width = 16 / 2,
  fig.height = 9 / 2,
  message = FALSE,
  warning = FALSE,
  cache = FALSE
)
```

@@

This package is designed to work for spatially mapped, repeat censused forests plots. The package has commands to fit models of tree growth based on neighborhood competition which can be used to estimate species-specific competition coefficients. The models fits can then be evaluated using a spatial cross-validation scheme to detect possible overfitting. Additionally these models can test whether the species identity of competitors matters using a permutation test-style shuffling of competitor identity (under the null hypothesis) and subsequently evaluating if model performance changes. See Allen and Kim (2020) [A permutation test and spatial cross-validation approach to assess models of interspecific competition between trees](https://doi.org/10.1371/journal.pone.0229930) for a full description.

@@

Here we provide an example with a small dataset.

```{r}
library(tidyverse)
library(forestecology)
library(sf)
library(sfheaders)
library(blockCV)
library(yardstick)
library(snakecase)
```

First we combine two example census data files into a single `tibble` and compute the growth of all surviving individuals.

@@

# Filter out resprouts
census_df2_ex_no_r <- census_df2_ex %>%
  filter(!str_detect(codes, "R"))

# Name of variable in tibble that uniquely identifies each stem:
id <- "ID"

# Merge both censuses and compute growth:
ex_growth_df <-
  compute_growth(census_df1_ex, census_df2_ex_no_r, id) %>%
  mutate(
    sp = to_any_case(sp),
    # Must be a factor
    sp = as.factor(sp)
  )
```

All growth models in the package assume that two individuals compete if they are less than a specified distance away. Here we set this distance in `max_dist` and put all individuals within this distance of a plot boundary in a buffer (since we do not have full information about their competitors).

```{r}
# Load study region:
data("ex_study_region")

# Set max dist
max_dist <- 1

# Add buffer
ex_growth_df <- ex_growth_df %>%
  add_buffer_variable(direction = "in", size = max_dist, region = ex_study_region)

@@

We then set blocks/folds for the spatial cross-validation scheme. We this both manually and using the [`blockCV`](https://github.com/rvalavi/blockCV) package. Note that here we define the geometric structure of the folds manually using `sfheaders::sf_polygon()`:

```{r}
fold1 <- rbind(c(0, 0), c(5, 0), c(5, 5), c(0, 5), c(0, 0))
fold2 <- rbind(c(5, 0), c(10, 0), c(10, 5), c(5, 5), c(5, 0))
blocks <- bind_rows(
  sf_polygon(fold1),
  sf_polygon(fold2)
) %>%
  mutate(foldID = c(1, 2))
```

Next we assign each point to folds using `blockCV::spatialBlock()`

```{r}
ex_cv_grid <- spatialBlock(
  speciesData = ex_growth_df,
  verbose = FALSE,
  k = 2,
  selection = "systematic",
  blocks = blocks,
  showBlocks = FALSE
)

@@

# Deliverable
ggplot() +
  geom_sf(data = ex_growth_df, aes(col = buffer, shape = foldID), size = 2) +
  geom_sf(data = blocks, fill = "transparent")

# Convert to sf object
ex_cv_grid_sf <- ex_cv_grid$blocks %>%
  st_as_sf()
```

We then create a `focal_v_comp` tibble which has a row for each competing pair of individuals (each pair of individuals within `max_dist` of one another):

```{r}
focal_vs_comp_ex <- ex_growth_df %>%
  create_focal_vs_comp(max_dist, cv_grid_sf = ex_cv_grid_sf, id = "ID")
```

With `focal_v_comp`, we can then run the growth model and get predictions from it. Here we run the model without cross validation. Posterior estimates can be plotted to give, for example, the lambda matrix of competition coefficients.

```{r}
# Fit model
posterior_param_ex <- focal_vs_comp_ex %>%
  fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)

# Get predicted dbh values values
predictions <- focal_vs_comp_ex %>%
  predict_bayesian_model(posterior_param = posterior_param_ex) %>%
  right_join(ex_growth_df, by = c("focal_ID" = "ID"))

# Compute RMSE of true vs predicted dbh values
predictions %>%
  rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)

# Plot posteriors
plot_ex <- posterior_param_ex %>%
  plot_posterior_parameters()
plot_ex[["lambda"]]
```

Here we repeat the process but with cross-validation. Note the increase in RMSE, reflecting the fact that our original estimate of model error was overly optimistic.

```{r}
# Fit model with cross-validation
ex_bw <- focal_vs_comp_ex %>%
  run_cv(max_dist = max_dist, cv_grid = ex_cv_grid) %>%
  right_join(ex_growth_df, by = c("focal_ID" = "ID"))

# Compute RMSE of true vs predicted dbh values
ex_bw %>%
  rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)
```
